A generalized definition of multidimensional item response theory parameters

### 2022-11-25 output/paper_draft/paper_M2PLM_parameters.Rmd
Changes definition of direction angles in test space, using the "norm" of the discrimination vector (instead of the equivalent square root of the inner product)
Deletes the TODO's in Mario's review (they are intended to be included in the Github review of the PR as comments.

Let's consider the standard basis $`r LS_STD_BASIS_EQ`$ in $`r LATENT_SPACE`$.
<!-- TODO: Incomplete in Mario's version, is it correct to say "in the latent space"? -->
We can write all $`r TRAIT_VEC_IN_LS`$ as a linear combination of each baseis.
Therefore, we have both representations: $`r TRAIT_VECTOR_DEF_BASIS_EQ`$, and
$`r TRAIT_VECTOR_DEF_STD_BASIS_EQ`$, where $`r TRAIT_COORDS_EQ`$ and
$`r TRAIT_COORDS_STD_EQ`$.

@@

The $`r DIM_INDEX`$-th column of $`r TRANSFORM_MATRIX`$ is given by the
coordinates of $`r BASIS_VECTOR_ANY`$ in basis $`r LS_STD_BASIS`$.
<!-- # TODO: For all k? -->
If $`r BASIS_VECTOR_ANY_TRANSF_EQ`$, then

@@

With the standard basis in $`r LATENT_SPACE`$ we have the usual scalar product<!-- (i.e. dot product) -->.
With basis $`r LS_BASIS`$ in $`r LATENT_SPACE`$ we define an inner product
$`r LS_BASIS_INNER_PROD`$, such that the norm of $`r TRAIT_VECTOR`$ is
invariant, i.e., $`r TRAIT_NORM_EQ`$. Therefore,

@@

The cosines between two vectors follow the usual definition,
<!-- # TODO: Is this true?? See Harman 1970 (eq. 4.43, p. 60 to check) -->

@@

\begin{equation}
  `r DIAG_MATRIX_INNER_PROD_EQ`.
\end{equation}
<!-- # TODO: I have changed here $D^{-1/2}$ by $D$; discuss with Mario -->

@@

and $`r TRANSFORM_MATRIX_ALT_EQ`$.
<!-- TODO: Comment with Mario (order of terms inverted here) -->

@@

Thus, the M2PL model is expressed in polar coordinates as
<!-- # TODO: Use the "module" symbol instead of the "norm"? -->

@@

Following the procedure outlined by @reckase_difficulty_1985
[@reckase_discriminating_1991],
we first compute the point of maximum slope by finding the root(s) of
the second derivative with respect to $`r TRAIT_NORM`$.
<!-- # TODO: Use the "module" symbol instead of the "norm"? -->

@@

To compute the direction of the maximum slope,
Reckase leverages on the property that the squared direction cosines sum up to 1
(see Equation 7 therein).
However, this assertion implicitly assumes orthogonal coordinates,
so it does not apply in general to $`r DIR_COS_VEC`$.
Instead, we may consider the direction cosine vector with the standard basis
$`r LS_STD_BASIS`$.<!-- # TODO: Mario wrote "\mathcal{B}", I think he refers to "\mathcal{E}", the standard basis -->
Now, applying Proposition 2 (with $`r ALT_MATRIX_AS_ID`$) we obtain

@@

Substituting $`r DIR_COS_VEC_STD`$
(Equation \@ref(eq:direction-cosine-vectors-eq)) and
$`r DISCR_VECTOR_STD`$ (Equation \@ref(eq:discr-vector-std))
by their respective expressions results in the
expression for the direction cosine vector of item $i$ in $`r LATENT_SPACE`$,
<!-- # TODO: with basis \mathcal{B}? -->

@@

which gives the direction from the origin to the point where
the slope is maximum.
Finally, to determine the signed distance $`r DISTANCE_PARAM`$
from the origin to that point,
we solve Equation \@ref(eq:M2PL-formula-polar) for $`r IRF_MAX_SLOPE_EQ`$ to get
$`r TRAIT_NORM_SOLVED_EQ`$. Using equation \@ref(eq:dir-cos-vec), then
<!-- # TODO: Verify that it's equation 20, not 18 -->

@@

\begin{equation}
  \begin{split}
    `r DIR_COS_ITEM_VEC_NORM_EQ` \\
    `r DISTANCE_PARAM_NORM_EQ`   \\
    `r SLOPE_MAX_PARAM_NORM_EQ`.
  \end{split}                             (\#eq:item-params-discrimination-norm)
\end{equation}
<!-- # TODO: Three equations collapsed into one altogether -->

It is clear that $`r INNER_PROD_INV_MAT_ELEMENT_EQ`$ were
$`r INNER_PROD_INV_MAT_ELEMENT`$ is the $`r INNER_PROD_INV_MAT_SUBINDEX`$-th
element of $`r INNER_PROD_MATRIX_INV`$.
<!-- # TODO: Error here? Mario means product of the basis elements? -->

Let $`r LS_STD_BASIS`$ be the standard basis of $`r TEST_SPACE`$.
<!-- # TODO: Can the standard basis of \Alpha be also \mathcal{E}? Otherwise, may I use \mathcal{E} instead? -->
Then, $`r BASIS_CHANGE_TS_EQ`$ is the change of basis matrix and
<!-- # TODO: Result pending of demonstration? -->
$`r DISCR_VECTOR_STD_MOD_EQ`$.

>>>>

Let's consider the standard basis $`r LS_STD_BASIS_EQ`$ in $`r LATENT_SPACE`$.
We can write all $`r TRAIT_VEC_IN_LS`$ as a linear combination of each baseis.
Therefore, we have both representations: $`r TRAIT_VECTOR_DEF_BASIS_EQ`$, and
$`r TRAIT_VECTOR_DEF_STD_BASIS_EQ`$, where $`r TRAIT_COORDS_EQ`$ and
$`r TRAIT_COORDS_STD_EQ`$.

@@

The $`r DIM_INDEX`$-th column of $`r TRANSFORM_MATRIX`$ is given by the
coordinates of $`r BASIS_VECTOR_ANY`$ in basis $`r LS_STD_BASIS`$.
If $`r BASIS_VECTOR_ANY_TRANSF_EQ`$, then

@@

With the standard basis in $`r LATENT_SPACE`$ we have the usual scalar product.
With basis $`r LS_BASIS`$ in $`r LATENT_SPACE`$ we define an inner product
$`r LS_BASIS_INNER_PROD`$, such that the norm of $`r TRAIT_VECTOR`$ is
invariant, i.e., $`r TRAIT_NORM_EQ`$. Therefore,

@@

The cosines between two vectors follow the usual definition,

@@

\begin{equation}
  `r DIAG_MATRIX_INNER_PROD_EQ`.
\end{equation}

@@

and $`r TRANSFORM_MATRIX_ALT_EQ`$.

@@

Thus, the M2PL model is expressed in polar coordinates as

@@

Following the procedure outlined by @reckase_difficulty_1985
[@reckase_discriminating_1991],
we first compute the point of maximum slope by finding the root(s) of
the second derivative with respect to $`r TRAIT_NORM`$.

@@

To compute the direction of the maximum slope,
Reckase leverages on the property that the squared direction cosines sum up to 1
(see Equation 7 therein).
However, this assertion implicitly assumes orthogonal coordinates,
so it does not apply in general to $`r DIR_COS_VEC`$.
Instead, we may consider the direction cosine vector with the standard basis
$`r LS_STD_BASIS`$.
Now, applying Proposition 2 (with $`r ALT_MATRIX_AS_ID`$) we obtain

@@

Substituting $`r DIR_COS_VEC_STD`$
(Equation \@ref(eq:direction-cosine-vectors-eq)) and
$`r DISCR_VECTOR_STD`$ (Equation \@ref(eq:discr-vector-std))
by their respective expressions results in the
expression for the direction cosine vector of item $i$ in $`r LATENT_SPACE`$,

@@

which gives the direction from the origin to the point where
the slope is maximum.
Finally, to determine the signed distance $`r DISTANCE_PARAM`$
from the origin to that point,
we solve Equation \@ref(eq:M2PL-formula-polar) for $`r IRF_MAX_SLOPE_EQ`$ to get
$`r TRAIT_NORM_SOLVED_EQ`$. Using equation \@ref(eq:dir-cos-vec), then

@@

\begin{equation}
  \begin{split}
    `r DIR_COS_ITEM_VEC_NORM_EQ` \\
    `r DISTANCE_PARAM_NORM_EQ`   \\
    `r SLOPE_MAX_PARAM_NORM_EQ`.
  \end{split}                             (\#eq:item-params-discrimination-norm)
\end{equation}

It is clear that $`r INNER_PROD_INV_MAT_ELEMENT_EQ`$ were
$`r INNER_PROD_INV_MAT_ELEMENT`$ is the $`r INNER_PROD_INV_MAT_SUBINDEX`$-th
element of $`r INNER_PROD_MATRIX_INV`$.

Let $`r LS_STD_BASIS`$ be the standard basis of $`r TEST_SPACE`$.
Then, $`r BASIS_CHANGE_TS_EQ`$ is the change of basis matrix and
$`r DISCR_VECTOR_STD_MOD_EQ`$.
